Elite individuals, insitutions, and economic growth accounting

### 2024-04-11 presentation/main.tex
actual vF

\begin{frame}{Research Question}{Importance and Relevance}
    \begin{block}{Does elite performance matters in education and human capital?}
        \begin{itemize}
            \item Rough measures of human capital such as school enrollment rate, are often used (e.g. Mankiw-Romer-Weil)
            \item Elite students and top university research may give countries a technology advantage
        \end{itemize}
    \end{block}

    \begin{block}{Implications for economic convergence}
        \begin{itemize}
            \item Richer, larger countries will likely have better research output and universities
            \item Technology and trade conflict between China-US have shown that ``hoarding'' technology may widen gap between nations
        \end{itemize}

@@

\begin{frame}{Rationale for variables}
    \begin{block}{IMO Scores}
        Indicator for a country's (and region's) ability to develop/identify pinncale STEM talent at high school level.
    \end{block}

@@

    \begin{block}{Percent in PISA 99th percentile}
        When controlling for average PISA math scores, this is a partial indicator for whether general excellence in academics is encouraged/necessary.
    \end{block}
\end{frame}

\section{Empirical Results}
\begin{frame}{Model specification}
    \small
    Let elite indicators be: $math99, ARWU, IMO$.
    For the sake of concision, let $E_{i,t}$ be a 1 by 4 matrix defined as:
    \[E_{i,t} = 
    \begin{bmatrix}
        math99_{i, t} & ARWU_{i, t} & IMO_{i, t}
    \end{bmatrix}
    \] for country/region $i$ in year $t$.
    For control variables, $I_{i, t}$ is the matrix of variables and $\alpha$ is coefficients.
    \begin{equation}
        Y_{i, t} = \beta_0 + \lambda E_{i, t} + \delta GDPpc \times E_{i,t} + \alpha I_{i, t} + T_t + C_i + \epsilon_{i, t}
    \end{equation}
    where $T,C$ represent time and entity dummies respectively and $Y_{i,t}$ be the GDP per capita growth in basis points for a country/region $i$ and year $t$.
    \textbf{Controls: school completion rates, GDP per capita, PISA average math score, democracy index}
\end{frame}

@@

\begin{frame}{PISA panel regression}{Without imputation}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{5pt}}lcccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{4}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-5}
                \\[-1.8ex] & \multicolumn{1}{c}{Model 1 (base)} & \multicolumn{1}{c}{Model 2} & \multicolumn{1}{c}{Model 3 (Time FE)} & \multicolumn{1}{c}{Model 4 (Time + Entity FE)}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & -31.228$^{*}$ & 5.985$^{}$ & 8.659$^{}$ & 25.421$^{}$ \\
                & (16.314) & (22.468) & (21.493) & (84.291) \\
                 IMO score per log population & 10.674$^{*}$ & 6.824$^{}$ & 12.529$^{}$ & 50.506$^{}$ \\
                & (6.363) & (9.440) & (8.631) & (32.187) \\
                 ARWU insitutions & -166.641$^{**}$ & -190.750$^{}$ & -249.208$^{**}$ & -454.384$^{}$ \\
                & (68.800) & (127.229) & (116.476) & (351.559) \\
                 Time Effects & No & No & Yes & Yes \\
                 Fixed Effects & No & No & No & Yes \\
                 Controls & No & Yes & Yes & Yes \\
                \hline \\[-1.8ex]
                 Observations & 440 & 112 & 112 & 112 \\
                 $R^2$ & 0.037 & 0.204 & 0.383 & 0.715 \\
                 Adjusted $R^2$ & 0.031 & 0.125 & 0.294 & 0.355 \\
                 Residual Std. Error & 445.700 (df=436) & 236.761 (df=101) & 212.750 (df=97) & 203.207 (df=49) \\
                 F Statistic & 5.616$^{***}$ (df=3; 436) & 2.587$^{***}$ (df=10; 101) & 4.294$^{***}$ (df=14; 97) & 1.987$^{***}$ (df=62; 49) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}
        \small
        \begin{itemize}
            \item Relationships are very different from imputed version
            \item Some countries are missing in this regression and some years as well
            \item School completion rates are only available for certain years and we have seen the relationship to be volatile year-to-year
        \end{itemize}
\end{frame}

@@

>>>>

\begin{frame}{Research Question}{Importance and Relevance}
    \begin{block}{Does elite performance matters in education and human capital?}
        \begin{itemize}
            \item Rough measures of human capital such as school enrollment rate, are often used (e.g. Mankiw-Romer-Weil)
            \item Elite students and top university research may give countries a human capital and/or research advantage
        \end{itemize}
    \end{block}

    \begin{block}{Implications for economic convergence}
        \begin{itemize}
            \item Richer, larger countries have better research output and universities
            \item Rise of outsourcing, technology sanctions indicate increasing importance of human capital
        \end{itemize}

@@

\begin{frame}{Rationale for variables}{What is defined as ``elite''?}
    \begin{block}{IMO Scores}
        Indicator for a country's (and region's) ability to develop and identify pinncale STEM talent at high school level. Requires students to take it seriously.
    \end{block}

@@

    \begin{block}{\% in PISA 99th percentile}
        Average scores are controlled for, so this indicates degree of right skew and wider academic culture.
    \end{block}
\end{frame}

\section{Empirical Results}
\begin{frame}{Model specification}
    \small
    Let elite indicators be: $math99, ARWU, IMO$.
    For the sake of concision, let $E_{i,t}$ be a 1 by 4 matrix defined as:
    \[E_{i,t} = 
    \begin{bmatrix}
        math99_{i, t} & ARWU_{i, t} & IMO_{i, t}
    \end{bmatrix}
    \] for country/region $i$ in year $t$.
    For control variables, $I_{i, t}$ is the matrix of variables and $\alpha$ is coefficients.
    \begin{equation}
        Y_{i, t} = \beta_0 + \lambda E_{i, t} + \alpha I_{i, t} + T_t + C_i + \epsilon_{i, t}
    \end{equation}
    where $T,C$ represent time and entity dummies respectively and $Y_{i,t}$ be the GDP per capita growth in basis points for a country/region $i$ and year $t$.
    \textbf{Controls: school completion rates, GDP per capita, PISA average math score, democracy index}
\end{frame}

@@

@@

\begin{frame}{Appendix: PISA panel regression}{Without imputation}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{5pt}}lcccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{4}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-5}
                \\[-1.8ex] & \multicolumn{1}{c}{Model 1 (base)} & \multicolumn{1}{c}{Model 2} & \multicolumn{1}{c}{Model 3 (Time FE)} & \multicolumn{1}{c}{Model 4 (Time + Entity FE)}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & -31.228$^{*}$ & 5.985$^{}$ & 8.659$^{}$ & 25.421$^{}$ \\
                & (16.314) & (22.468) & (21.493) & (84.291) \\
                 IMO score per log population & 10.674$^{*}$ & 6.824$^{}$ & 12.529$^{}$ & 50.506$^{}$ \\
                & (6.363) & (9.440) & (8.631) & (32.187) \\
                 ARWU insitutions & -166.641$^{**}$ & -190.750$^{}$ & -249.208$^{**}$ & -454.384$^{}$ \\
                & (68.800) & (127.229) & (116.476) & (351.559) \\
                 Time Effects & No & No & Yes & Yes \\
                 Fixed Effects & No & No & No & Yes \\
                 Controls & No & Yes & Yes & Yes \\
                \hline \\[-1.8ex]
                 Observations & 440 & 112 & 112 & 112 \\
                 $R^2$ & 0.037 & 0.204 & 0.383 & 0.715 \\
                 Adjusted $R^2$ & 0.031 & 0.125 & 0.294 & 0.355 \\
                 Residual Std. Error & 445.700 (df=436) & 236.761 (df=101) & 212.750 (df=97) & 203.207 (df=49) \\
                 F Statistic & 5.616$^{***}$ (df=3; 436) & 2.587$^{***}$ (df=10; 101) & 4.294$^{***}$ (df=14; 97) & 1.987$^{***}$ (df=62; 49) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}
        \small
        \begin{itemize}
            \item Relationships are very different from imputed version
            \item Some countries are missing in this regression and some years as well
            \item School completion rates are only available for certain years and we have seen the relationship to be volatile year-to-year
        \end{itemize}
\end{frame}
